---
title: "Marta"
author: "Marta"
date: "11/5/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

understanding packages

```{r packages}
install.packages("expm")
install.packages("gee")
install.packages("igraph")
install.packages("ape")
install.packages("fastmatch")
install.packages("quadprog")
install.packages("phangorn")
install.packages("phytools")
install.packages("geiger")
```

```{r}
library(ape)
library(fastmatch)
library(quadprog)
library(phangorn)
library(phytools)
library(geiger)
```

```{r nex}
nex.tree<-read.nexus.data(file="matrix_from_app2_xl_file.nex")
nexus2DNAbin(nex.tree)
plot.phylo(nex.tree)
```
ahhhhhhhhhhhhhhhhhhhhhhhhhhhhhhhhhhhhhhhhhhh ok i'm sleeping now but i think the answer is converting this xl file to readable tre
```{r}
as.phylo.formula(~Order/Family/Genus/Species, data =  (file="matrix_from_app2_xl_file.nex"))
```

```{r}
nex2.tree<-read.newick("matrix_from_app2_xl_file.nex")
plotTree(nex2.tree)
```

```{r}
install.packages("TreeTools")
library(TreeTools)
```
```{r}
as_phydat <- ReadAsPhyDat("matrix_from_app2_xl_file.nex")
as_phydat
summary(as_phydat)
raw_data <- ReadCharacters("matrix_from_app2_xl_file.nex")
my_data <- MatrixToPhyDat(raw_data)#i think these are important
```

```{r}
read.nexus.data("matrix_from_app2_xl_file.nex")
```

```{r}
library(help = geiger)
library(geiger)
```
```{r}
string <- "((A_alala_negra, A_corcyra_aretina, A_tracta, A_pithys, A_donysa_donysa), ((A_serpacelerio, A_seriphiaaquillia, A_seriphiatherasia), (A_melonaleucocoma, A_salmoneus_colada, A_cytherea_cytherea, A_cythereadaguana, A_epioneagilla, A_etheldaethelda, A_thessaliathessalia, A_iphicleolathessalita, A_iphiclusiphiclus, A_shuara, A_plesaurephliassa, A_basiloides, A_atticaattica, A_leucerioides, A_saundersiisaundersii, A_boreasboreas, A_rothschildi, A_sichaeus, A_cocalacocala, A_leucophthalmairminella, A_irminatumida, A_pollina, A_lycoriaslara, A_lycoriasspruceana, A_erotia_erotia, A_mesentina, A_phylaca_pseudaethalia, A_capucinus_capucinus, A_heraclea_heraclea, A_naxia_naxia, A_justina_valentina, A_olynthia, A_jordani, A_zinazina, A_zinairma, A_delinita_delinita, A_boeotia_boeotia, A_maleaaethalia)));"

morph.tree<-read.tree(text=string)
plot(morph.tree,no.margin=TRUE,edge.width=2)
```
```{r}
#attempt
#text='((A:1, B:1):2, ((C:1, D:1):2, (E:1, F:1):2):4);')
mol.tree<-read.nexus(file="NJst.tre")
mol.tree

myTree <- ape::read.tree(text= '((A_alala_negra:1, A_corcyra_aretina:1, A_tracta:1, A_pithys:1, A_donysa_donysa:1):3, ((A_serpacelerio:1, A_seriphiaaquillia:1, A_seriphiatherasia:1):2, (A_melonaleucocoma:1, A_salmoneus_colada:1, A_cytherea_cytherea:1, A_cythereadaguana:1, A_epioneagilla:1, A_etheldaethelda:1, A_thessaliathessalia:1, A_iphicleolathessalita:1, A_iphiclusiphiclus:1, A_shuara:1, A_plesaurephliassa:1, A_basiloides:1, A_atticaattica:1, A_leucerioides:1, A_saundersiisaundersii:1, A_boreasboreas:1, A_rothschildi:1, A_sichaeus:1, A_cocalacocala:1, A_leucophthalmairminella:1, A_irminatumida:1, A_pollina:1, A_lycoriaslara:1, A_lycoriasspruceana:1, A_erotia_erotia:1, A_mesentina:1, A_phylaca_pseudaethalia:1, A_capucinus_capucinus:1, A_heraclea_heraclea:1, A_naxia_naxia:1, A_justina_valentina:1, A_olynthia:1, A_jordani:1, A_zinazina:1, A_zinairma:1, A_delinita_delinita:1, A_boeotia_boeotia:1, A_maleaaethalia:1):2):3):4;')
plot(myTree)

comparePhylo(myTree, mol.tree, plot = FALSE, force.rooted = FALSE, use.edge.length = FALSE)

comparePhylo(myTree, mol.tree, plot = FALSE, force.rooted = TRUE,
             use.edge.length = FALSE)
```

Comparison Final Essay
In the end the major take away is that, while useful, tree comparison between two different types of trees can be incredibly difficult. Morphological trees have their benefits but often do not have the material necessary to stand up to arduous statistical analysis, as we have seen here today. As genomic research advances, we have more access to molecular data 